Measuring Interaction Design \\before Building the System: a Model-Based Approach

### 2016-01-09 paper.tex
submitted abstract to eics16
changed the title, extended the abstract, written the contribution.

\title{Measuring Interaction Design}

@@

@@

  We present a tool (MIGTool) that transforms models of a user
  interface into a graph, upon which usage scenarios can be easily
  specified, and used to compute possible interaction paths. Metrics
  based on possible paths, with or without user errors, can be easily
  computed.  For example, when applying them to compare mail
  applications, we show that Gmail has 3 times more shortest routes,
  has 2x more routes that include a single user error, has routes with
  13\% fewer steps, but optimal routes have the smallest probability
  to be chosen.

>>>>

\title{Measuring Interaction Design \\before Building the System: a
  Model-Based Approach}

@@

% \author{Giorgio Brajnik\inst{1,2} \and Simon Harper\inst{1}}
% \institute{\textsuperscript{1}Computer Science School\\
% University of Manchester, UK\\
% \email{simon.harper@manchester.ac.uk}\\
% and\\
% \inst{2}Dipartimento di Matematica e Informatica\\
% Universit\`a di Udine, Italy\\
% \email{brajnik@uniud.it}
% }

@@

  We present a tool (MIGTool) that transforms models of the behavior
  of a user interface into a graph, upon which usage scenarios can be
  easily specified, and used by MIGTool to compute possible
  interaction paths. Metrics based on possible paths, with or without
  user navigation errors, can then be computed.  For example, when
  analyzing four mail applications, we show that Gmail compared to
  others has 3 times more shortest routes, has 2x more routes that
  include a single user error, has routes with 13\% fewer steps, but
  has also optimal routes with the smallest probability to be chosen.

  Without MIGTool, this kind of analysis could only be done after
  building some prototype of the system, and then only for specific
  scenarios by manually tracing user actions and relative changes to
  the screens. With MIGTool the exploration of suitability of a design
  with respect to different scenarios, or comparison of different
  design alternatives, can be done after providing only a partial
  specification of the user interface behavior.

  This is made possible by the ability to associate scenarios steps to
  required user actions as defined in the model, by an efficient
  strategy to identify complete execution traces that users can
  follow, and by computing a range of diverse metrics on these
  results. The agile modeling approach that was followed implies that
  relatively little work is needed upfront.

  CONTRIBUTION (asked when submitting the abstract)

  There are three contributions.

  First, we provide means to concisely ground usage scenarios on the
  basis of actions specified in models. This mechanism is general
  enough that can be used also to link tasks or use cases to actual
  behavior of a user interface.

  Second, MIGTool transforms models and scenarios into a graph of
  possible execution paths. Thanks to the special-purpose language
  that is provided to specify scenarios, MIGTool is quite efficient,
  with a time complexiy that is linear with the number of steps in a
  scenario and the number of edges in the graph. We describe the
  language and the algorithm.

  Third, we provide definition of a handful of metrics that are
  objective, precise and that bear upon usability of the resulting
  user interface. They are very general, can be used for most kinds of
  user interfaces that can be modeled with discrete events, and are
  platform independent.

%% EICS submission: #122
%% https://precisionconference.com/~eics16a/updatePaperResponse?userRef=CILXDODXPNF0VESOAFMMC8I16ZO5FR5J&paperNumber=112&emailSent=&error=
